---
title: "Simulations of permutation tests"
author: "Ruiting Tong"
date: "1/29/2023"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.height = 5, fig.width = 10)
```

Simulate 150 dates. Suppose there are only 3 days in a week. Sunday: +0 to duration; Monday: +5 to duration; Tuesday: +15 to duration. The decrease in duration: Every Sunday - 0.02, every Monday - 0.04, every Tuesday - 0.1. 

Noise: Normal mean 0 sd 5

Methods:

1. Ungrouped: Spearman Correlation Test
2. Grouped: Divide data into 3 groups based on weekdays and permute within each group. Compare the Spearman corrleation test statistic with its empirical distribution
3. Grouped: Divide data into 3 groups based on weekdays and compute the test statistic by adding the test statistics of Spearman correlation test run in each group together. Compare this statistic with its emprical distribution 

Number of p-values computed for each method: N = 500

```{r}
dates = 150
N = 500
p_values_spearman = numeric(N)

test_stats = numeric(N)
p_values_groupShuffle = numeric(N)
p_values_grouped = numeric(N)

decrease = ((1:dates)%%3==1)*(1:dates+2)*0.02/3 + 
  ((1:dates)%%3==2)*(1:dates+1)*0.04/3 + 
  ((1:dates)%%3==0)*(1:dates)*0.1/3

# Toy example with little noise
jpeg("Durations.jpg", width = 960, height = 480)
par(mfrow=c(1,2))

durations = rep(c(0,5,15), dates/3) + (500-decrease)
plot(durations, col=c(1:3))
legend("topright", legend = c("Sunday", "Monday", "Tuesday"), 
       pch = 1, col=c(1:3))

# Toy example with noise sd = 5
durations = rep(c(0,5,15), dates/3) + (500-decrease) + rnorm(dates, 0, 5) 
plot(durations, col=c(1:3))
legend("topright", legend = c("Sunday", "Monday", "Tuesday"), 
       pch = 1, col=c(1:3))

for(i in 1:N){
  durations = rep(c(0,5,15), dates/3) + (500-decrease) + rnorm(dates, 0, 5) 
  result = cor.test(durations, 1:dates, "less", "spearman")
  p_values_spearman[i] = result$p.value
  test_stat_spearman = result$statistic

  
  test_stat_shuffle = numeric(1000)
  for(j in 1:1000){
    shuffled = numeric(dates)
    Sun = durations[3*(1:50)-2]  
    Mon = durations[3*(1:50)-1]
    Tue = durations[3*(1:50)]
    shuffled[3*(1:50)-2] = Sun[sample(50)]
    shuffled[3*(1:50)-1] = Mon[sample(50)]
    shuffled[3*(1:50)] = Tue[sample(50)]
    
    test_stat_shuffle[j] = sum((rank(shuffled) - 1:dates)^2) 
  }
  
  p_values_groupShuffle[i] = 1- 
    sum((test_stat_spearman > test_stat_shuffle))/1000
  
  
  test_stats[i] = sum((rank(durations[seq(1,148, by = 3)]) - 1:50)^2) + 
  sum((rank(durations[seq(2,149, by = 3)]) - 1:50)^2) +
  sum((rank(durations[seq(3,150, by = 3)]) - 1:50)^2)
}
```

```{r}
B = 10000
test_stats_null = numeric(B)
for(i in 1:B){
  duration_rks = sample(150)
  test_stats_null[i] = 
    sum((rank(duration_rks[seq(1,dates-2, by = 3)]) - 1:50)^2) + 
    sum((rank(duration_rks[seq(2,dates-1, by = 3)]) - 1:50)^2) +
    sum((rank(duration_rks[seq(3,dates, by = 3)]) - 1:50)^2)
}

hist(test_stats_null)
```

```{r}
for(i in 1:N){
  p_values_grouped[i] = 1- sum(test_stats[i] > test_stats_null)/B
}


jpeg("Two-Methods.jpg", height = 640, width = 960)
par(mfrow = c(2,1))
hist(p_values_spearman, 
     ylab = "frequency", xlab = "p-values",
     main = "Histogram of p-values with no grouping")
hist(p_values_groupShuffle, 
     ylab = "frequency", xlab = "p-values",
     main = "Historgram of p-values by permutations within groups.")

```




